opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-12-05 chapters/03_existing_datasets.tex
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;

\begin{table}[!htb]
\centering
\caption{Overview of Experimental Powder Diffraction Databases}
\label{tab:exp_databases}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Name}                          & \textbf{No. of Patterns} & \textbf{Year established} \\
\midrule
Linus Pauling file                 & 21,700                  & 2002        \\
Powder Diffraction File (PDF)  \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.}         & 20,800                  & 1941                      \\
RRUFF                                  & 1290                    & 2006                      \\
Crystallography Open Database (COD)    & 1063                    & 2003                      \\
PowBase                                & 169                     & 1999                      \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{table}[!htb]
\centering
\caption{Overview of Experimental Powder Diffraction Databases}
\label{tab:exp_databases}
\scalebox{0.85}{
\begin{tabular}{@{}cccccccc@{}}
\toprule
\textbf{Name} & \textbf{No. of Patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg} & \textbf{Lattice} & \textbf{Unit Cell} & \textbf{Year est.} \\
\midrule
Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
Powder Diffraction File \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \cmark & 1941 \\
RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
Crystallography Open Database          & 1063                      & \cmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2003            \\
PowBase                                & 169                       & \cmark          & \xmark        & \xmark        & \xmark        & \xmark          & 1999            \\
\bottomrule
\end{tabular}}
\end{table}

### 2024-12-05 chapters/04_our_dataset.tex
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;

\begin{table}[!htb]
\centering
\caption{opXRD database entries}
\label{tab:merged}
\begin{tabular}{@{}clcccccl@{}}
\toprule
Institution & No. Patterns & Comp. & Spg & Lattice & Unit Cell & Research Project \\ 
\midrule
CNRS        & \qquad 895          & \cmark   & \cmark   & \cmark   & \cmark   & Extracted from the COD                 \\ 
HKUST(GZ)   & \qquad 21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
USC         & \qquad 338          & \cmark   & \cmark   & \cmark   & \xmark   & Study of CuNi and CuAl alloys          \\ 
EMPA        & \qquad 790          & \cmark   & \cmark   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
INT         & \qquad 19780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
IKFT        & \qquad 64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
LBNL        & \qquad n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure*}[!ht]
    \centering
    \missingfigure{} 
    \caption{Statistics, histograms, etc. of our dataset.}
    \label{fig:statistics}
\end{figure*}

%% Paragraph by FX Coudert and Arthur Hardiagon
\subsubsection*{Chimie ParisTech, PSL University}

Experimental pXRD data was extracted from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}. The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries. The data in the COD is placed in the public domain and licensed under the CC0 License. Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies. These 5432 structures only account for 1\% of the total COD database, but this is to be expected since most crystal structures are resolved from single-crystal diffraction. Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself. We could extract raw experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data. \\

>>>>

\begin{table}[!htb]
\centering
\caption{opXRD database entries}
\label{tab:merged}
\scalebox{0.92}{
\begin{tabular}{@{}clcccccl@{}}
\toprule
Institution & No. Patterns & Comp. & Spg & Lattice & Unit Cell & Research Project \\ 
\midrule
CNRS        & \qquad 1052          & \cmark   & \cmark   & \cmark   & \cmark   & Extracted from the COD                 \\ 
HKUST(GZ)   & \qquad 21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
USC         & \qquad 338          & \cmark   & \cmark   & \cmark   & \xmark   & Study of CuNi and CuAl alloys          \\ 
EMPA        & \qquad 790          & \cmark   & \cmark   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
INT         & \qquad 19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
IKFT        & \qquad 64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
LBNL        & \qquad n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
\bottomrule
\end{tabular}}
\end{table}

%TODO:
% The three datasets shown here are (1): USC CuNi, (2): USC CuAl, (3): HKUST
% They certainly form clusters, but we have to think about how to interpret these results. It might still represent background/angle range more than anything else more interesting like material class.
% Also if we add a larger dataset like the INT data the variance is so huge that all these three more narrow datasets effectively collapse onto a tiny region (see below)
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\linewidth]{figures/pca_scatter_smol.png}
    \caption{Two component PCA scatter plot showing (1): USC CuNi, (2): USC CuAl, (3): HKUST data2}
    \label{fig:enter-label}
\end{figure}

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\linewidth]{figures/pca_scatter.png}
    \caption{Two component PCA scatter plot including all opXRD data}
    \label{fig:enter-label}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\linewidth]{figures/quantities_hist.png}
    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value and final angle value for all data}
    \label{fig:enter-label}
\end{figure}

\pagebreak
%% Paragraph by FX Coudert and Arthur Hardiagon
\subsubsection*{Chimie ParisTech, PSL University}

Experimental pXRD data was extracted from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}. The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries. The data in the COD is placed in the public domain and licensed under the CC0 License. Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies. These 5432 structures only account for 1\% of the total COD database, but this is to be expected since most crystal structures are resolved from single-crystal diffraction. Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself. We could extract raw experimental pXRD patterns from 1052 files in total, after curation of a small number of files with clearly invalid data. \\

### 2024-12-05 chapters/05_how_to_use.tex
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;

The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14254271}.} and can be downloaded by any user without any barriers or restrictions. The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements. Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series. \\

@@

Our tool further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors.
We provide the following Jupyter notebook that showcases the functionality of the library in more detail: <Link to github Jupyter notebook on Colab>

%The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method. The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.

>>>>

The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14278656}.} and can be downloaded by any user without any barriers or restrictions. The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements. Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series. \\

@@

Our library further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors. We provide a Jupyter Notebook\footnote{This notebook can be found here: \url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb}.} that showcases these functionalities in more detail. This notebook also illustrates how to interface with the OpXRD database through Python.
